---
title: "Comparison of fixed-effects and mixed-effects models"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)

library(tidyverse)
library(ggeffects)
library(lme4)
library(easystats)
library(boot)

```

# Table S1. Best fit bivariate models for termite discovery versus key spatial and climatic variables

Fixed- and mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv")  %>%
  mutate(abs_lat = abs(lat),
         date_diff_years = date_diff/365) %>%
  filter(trt %in% c("T") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara")

```

### Temperature

Fixed only:

```{r}

modtemp<-glm(termite_exposure ~ temp,
            data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))

model_parameters(modtemp,exponentiate = TRUE)

```

Mixed:

```{r}

modtemp<-glmer(termite_exposure ~ temp + (1|site),
            data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))

model_parameters(modtemp,exponentiate = TRUE, effects='fixed')

```

### Precipitation

Fixed only:

```{r}

modpre<-glm(termite_exposure~prec,
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modpre,exponentiate = TRUE)

```

Mixed:

```{r}

modpre<-glmer(termite_exposure~prec + (1|site),
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modpre,exponentiate = TRUE, effects='fixed')

```

### Absolute latitude

Fixed only:

```{r}

modlat<-glm(termite_exposure~abs_lat,
             data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modlat,exponentiate = TRUE)

```

Mixed:

```{r}

modlat<-glmer(termite_exposure~abs_lat + (1|site),
             data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modlat,exponentiate = TRUE, effects='fixed')

```

### Elevation

Fixed only:

```{r}

modalt<-glm(termite_exposure~alt,
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modalt,exponentiate = TRUE)

```

Mixed:

```{r}

modalt<-glmer(termite_exposure~alt + (1|site),
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modalt,exponentiate = TRUE, effects='fixed')

```



# Table S2. Best fit bivariate models for decay (k) versus key spatial and climatic variables

Fixed-effects models are fitted after calculating site-level means for decay. Mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv") %>%
  filter(trt %in% c("T", "C") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara") %>%
  mutate(date_diff_years = date_diff/365, 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))


#get site level decay
mm %>% 
  group_by(site, termite_exposure) %>% 
  summarise(temp = temp[1], 
            prec = prec[1], 
            forestcover100 = forestcover100[1], 
            alt = alt[1],
            date_diff_years=date_diff_years[1],
            lat = lat[1],
            N_pc = N_pc[1],
            C_pc = C_pc[1],
            n = n(), 
            k_value= mean(k_value),
            #kval = log(k_value),
            discovered = mean(termite_exposure)) %>%
  ungroup() %>% 
  mutate(termite_exposure = factor(termite_exposure), 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))->mmm

```

### Absolute latitude -- termite decay

Fixed only:

```{r}

modlatt<-lm(k_value_tr ~abs_lat,
            data=filter(mmm,termite_exposure ==1))
model_parameters(modlatt)

```

Mixed:

```{r}

modlatt<-lmer(k_value_tr ~abs_lat + (1|site),
            data=filter(mm,termite_exposure ==1))
model_parameters(modlatt, effects='fixed')

```

### Absolute latitude -- microbial decay

Fixed only:

```{r}

modlatm<-lm(k_value_tr ~abs_lat,
            data=filter(mmm,termite_exposure ==0))
model_parameters(modlatm)

```

Mixed:

```{r}

modlatm<-lmer(k_value_tr ~abs_lat + (1|site),
            data=filter(mm,termite_exposure ==0))
model_parameters(modlatm, effects='fixed')

```


### Elevation -- termite decay

Fixed only:

```{r}

modaltt<-lm(k_value_tr ~alt,
            data=filter(mmm,termite_exposure ==1))
model_parameters(modaltt)

```

Mixed:

```{r}

modaltt<-lmer(k_value_tr ~alt + (1|site),
            data=filter(mm,termite_exposure ==1))
model_parameters(modaltt, effects='fixed')

```


### Elevation -- microbial decay

Fixed only:

```{r}

modaltm<-lm(k_value_tr ~alt,
            data=filter(mmm,termite_exposure ==0))
model_parameters(modaltm)

```

Mixed:

```{r}

modaltm<-lmer(k_value_tr ~alt + (1|site),
            data=filter(mm,termite_exposure ==0))
model_parameters(modaltm, effects='fixed')

```

### Temperature -- termite decay

Fixed only:

```{r}

modtempt<-lm(k_value_tr ~temp,
             data=filter(mmm,termite_exposure ==1))
model_parameters(modtempt)

```

Mixed:

```{r}

modtempt<-lmer(k_value_tr ~temp + (1|site),
             data=filter(mm,termite_exposure ==1))
model_parameters(modtempt, effects='fixed')

```


### Temperature -- microbial decay

Fixed only:

```{r}

modtempm<-lm(k_value_tr ~temp,
             data=filter(mmm,termite_exposure ==0))
model_parameters(modtempm)

```

Mixed:

```{r}

modtempm<-lmer(k_value_tr ~temp + (1|site),
             data=filter(mm,termite_exposure ==0))
model_parameters(modtempm, effects='fixed')

```


### Precipitation -- termite decay

Fixed only:

```{r}

modprect<-lm(k_value_tr ~prec,
             data=filter(mmm,termite_exposure ==1))
model_parameters(modprect)

```

Mixed:

```{r}

modprect<-lmer(k_value_tr ~prec + (1|site),
             data=filter(mm,termite_exposure ==1))
model_parameters(modprect, effects='fixed')

```


### Precipitation -- microbial decay

Fixed only:

```{r}

modprecm<-lm(k_value_tr ~prec,
             data=filter(mmm,termite_exposure ==0))
model_parameters(modprecm)

```

Mixed:

```{r}

modprecm<-lmer(k_value_tr ~prec + (1|site),
             data=filter(mm,termite_exposure ==0))
model_parameters(modprecm, effects='fixed')

```


# Table S3. Best fit multivariate model for probably of termite discovery versus climatic sensitivities

Fixed- and mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv")  %>%
  mutate(abs_lat = abs(lat),
         date_diff_years = date_diff/365) %>%
  filter(trt %in% c("T") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara")

```

Fixed only: 

```{r}

modinteraction<-glm(termite_exposure ~ prec*temp,
                    data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))
model_parameters(modinteraction,exponentiate = TRUE)

```

Mixed:

```{r}

modinteraction<-glmer(termite_exposure ~ prec*temp + (1|site),
                    data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))
model_parameters(modinteraction,exponentiate = TRUE, effects='fixed')

```


# Table S4. Best fit multivariate model for microbe (termite undiscovered) wood decay (k) versus climatic sensitivities

Fixed-effects models are fitted after calculating site-level means for decay and weighting sites by the number of observations in each category. Mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv") %>%
  filter(trt %in% c("T", "C") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara") %>%
  mutate(date_diff_years = date_diff/365, 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))


#get site level decay
mm %>% 
  group_by(site, termite_exposure) %>% 
  summarise(temp = temp[1], 
            prec = prec[1], 
            forestcover100 = forestcover100[1], 
            alt = alt[1],
            date_diff_years=date_diff_years[1],
            lat = lat[1],
            N_pc = N_pc[1],
            C_pc = C_pc[1],
            n = n(), 
            k_value= mean(k_value),
            #kval = log(k_value),
            discovered = mean(termite_exposure)) %>%
  ungroup() %>% 
  mutate(termite_exposure = factor(termite_exposure), 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))->mmm

```

Fixed only: 

```{r}

m1w <- lm(k_value_tr ~ temp*prec_m ,
          data=filter(mmm,termite_exposure ==0), weights=n)
model_parameters(m1w)

```

Mixed:

```{r}

m1w <- lmer(k_value_tr ~ temp*prec_m + (1|site),
          data=filter(mm,termite_exposure ==0))
model_parameters(m1w, effects='fixed')

```

# Table S5. Best fit multivariate model for microbe + termite (termite discovered) wood decay (k) versus climatic sensitivities

Fixed-effects models are fitted after calculating site-level means for decay and weighting sites by the number of observations in each category. Mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv") %>%
  filter(trt %in% c("T", "C") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara") %>%
  mutate(date_diff_years = date_diff/365, 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))


#get site level decay
mm %>% 
  group_by(site, termite_exposure) %>% 
  summarise(temp = temp[1], 
            prec = prec[1], 
            forestcover100 = forestcover100[1], 
            alt = alt[1],
            date_diff_years=date_diff_years[1],
            lat = lat[1],
            N_pc = N_pc[1],
            C_pc = C_pc[1],
            n = n(), 
            k_value= mean(k_value),
            #kval = log(k_value),
            discovered = mean(termite_exposure)) %>%
  ungroup() %>% 
  mutate(termite_exposure = factor(termite_exposure), 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))->mmm

```

Fixed only: 

```{r}

m2w <- lm(k_value_tr ~ temp*prec_m ,
          data=filter(mmm,termite_exposure ==1), weights=n)
model_parameters(m2w)

```

Mixed:

```{r}

m2w <- lmer(k_value_tr ~ temp*prec_m + (1|site),
          data=filter(mm,termite_exposure ==1))
model_parameters(m2w, effects='fixed')

```


# Table S6. Best fit bivariate models for termite discovery versus key spatial and climatic variables and wood chemistry

Fixed- and mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv")  %>%
  mutate(abs_lat = abs(lat),
         date_diff_years = date_diff/365) %>%
  filter(trt %in% c("T") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara")

```

### Temperature

Fixed only:

```{r}

mod_temp_NC<-glm(termite_exposure ~ temp+N_pc+C_pc,
              data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))
model_parameters(mod_temp_NC,exponentiate = TRUE)

```

Mixed:

```{r}

modtemp_NC<-glmer(termite_exposure ~ temp+N_pc+C_pc + (1|site),
              data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))
model_parameters(modtemp_NC,exponentiate = TRUE, effects='fixed')

```

### Precipitation

Fixed only:

```{r}

modpre_NC<-glm(termite_exposure~prec+N_pc+C_pc,
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modpre_NC,exponentiate = TRUE)

```

Mixed:

```{r}

modpre_NC<-glmer(termite_exposure~prec+N_pc+C_pc + (1|site),
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modpre_NC,exponentiate = TRUE, effects='fixed')

```

### Absolute latitude

Fixed only:

```{r}

modlat_NP<-glm(termite_exposure~abs_lat+N_pc+C_pc,
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modlat_NP,exponentiate = TRUE)

```

Mixed:

```{r}

modlat_NP<-glmer(termite_exposure~abs_lat+N_pc+C_pc + (1|site),
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modlat_NP,exponentiate = TRUE, effects='fixed')

```

### Elevation

Fixed only:

```{r}

modalt_NC<-glm(termite_exposure~alt+N_pc+C_pc,
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modalt_NC,exponentiate = TRUE)

```

Mixed:

```{r}

modalt_NC<-glmer(termite_exposure~alt+N_pc+C_pc + (1|site),
            data=mm,offset=date_diff_years,family = binomial(link = "logit"))
model_parameters(modalt_NC,exponentiate = TRUE, effects='fixed')

```


# Table S7. Best fit bivariate models for decay (k) versus key spatial versus climatic variables and wood chemistry

Fixed-effects models are fitted after calculating site-level means for decay. Mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv") %>%
  filter(trt %in% c("T", "C") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara") %>%
  mutate(date_diff_years = date_diff/365, 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))


#get site level decay
mm %>% 
  group_by(site, termite_exposure) %>% 
  summarise(temp = temp[1], 
            prec = prec[1], 
            forestcover100 = forestcover100[1], 
            alt = alt[1],
            date_diff_years=date_diff_years[1],
            lat = lat[1],
            N_pc = N_pc[1],
            C_pc = C_pc[1],
            n = n(), 
            k_value= mean(k_value),
            #kval = log(k_value),
            discovered = mean(termite_exposure)) %>%
  ungroup() %>% 
  mutate(termite_exposure = factor(termite_exposure), 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))->mmm

#Termite decay across elevation

#Microbial decay across elevation
modaltm_NC<-lm(k_value_tr ~alt+N_pc+C_pc,
            data=filter(mmm,termite_exposure ==0)
)

#Termite decay across temperature

#Microbial decay across temperature
modtempm_NC<-lm(k_value_tr ~temp+N_pc+C_pc,
             data=filter(mmm,termite_exposure ==0)
)



```

### Absolute latitude -- termite decay

Fixed only:

```{r}

modlatt_NC<-lm(k_value_tr ~abs_lat+N_pc+C_pc,
            data=filter(mmm,termite_exposure ==1)
)
model_parameters(modlatt_NC)

```

Mixed:

```{r}

modlatt_NC<-lmer(k_value_tr ~abs_lat+N_pc+C_pc + (1|site),
            data=filter(mm,termite_exposure ==1)
)
model_parameters(modlatt_NC, effects='fixed')

```

### Absolute latitude -- microbial decay

Fixed only:

```{r}

modlatm_NC<-lm(k_value_tr ~abs_lat+N_pc+C_pc,
            data=filter(mmm,termite_exposure ==0)
)
model_parameters(modlatm_NC)

```

Mixed:

```{r}

modlatm_NC<-lmer(k_value_tr ~abs_lat+N_pc+C_pc + (1|site),
            data=filter(mm,termite_exposure ==0)
)
model_parameters(modlatm_NC, effects='fixed')

```


### Elevation -- termite decay

Fixed only:

```{r}

modaltt_NC<-lm(k_value_tr ~alt+N_pc+C_pc,
            data=filter(mmm,termite_exposure ==1)
)
model_parameters(modaltt_NC)

```

Mixed:

```{r}

modaltt_NC<-lmer(k_value_tr ~alt+N_pc+C_pc + (1|site),
            data=filter(mm,termite_exposure ==1)
)
model_parameters(modaltt_NC, effects='fixed')

```


### Elevation -- microbial decay

Fixed only:

```{r}

modaltm_NC<-lm(k_value_tr ~alt+N_pc+C_pc,
            data=filter(mmm,termite_exposure ==0)
)
model_parameters(modaltm_NC)

```

Mixed:

```{r}

modaltm_NC<-lmer(k_value_tr ~alt+N_pc+C_pc + (1|site),
            data=filter(mm,termite_exposure ==0)
)
model_parameters(modaltm_NC, effects='fixed')

```


### Temperature -- termite decay

Fixed only:

```{r}

modtempt_NC<-lm(k_value_tr ~temp+N_pc+C_pc,
             data=filter(mmm,termite_exposure ==1)
)
model_parameters(modtempt_NC)

```

Mixed:

```{r}

modtempt_NC<-lmer(k_value_tr ~temp+N_pc+C_pc + (1|site),
             data=filter(mm,termite_exposure ==1)
)
model_parameters(modtempt_NC, effects='fixed')

```


### Temperature -- microbial decay

Fixed only:

```{r}

modtempm_NC<-lm(k_value_tr ~temp+N_pc+C_pc,
             data=filter(mmm,termite_exposure ==0)
)
model_parameters(modtempm_NC)

```

Mixed:

```{r}

modtempm_NC<-lmer(k_value_tr ~temp+N_pc+C_pc + (1|site),
             data=filter(mm,termite_exposure ==0)
)
model_parameters(modtempm_NC, effects='fixed')

```


### Precipitation -- termite decay

Fixed only:

```{r}

modprect_NC<-lm(k_value_tr ~prec+N_pc+C_pc,
             data=filter(mmm,termite_exposure ==1)
)
model_parameters(modprect_NC)

```

Mixed:

```{r}

modprect_NC<-lmer(k_value_tr ~prec+N_pc+C_pc + (1|site),
             data=filter(mm,termite_exposure ==1)
)
model_parameters(modprect_NC, effects='fixed')

```


### Precipitation -- microbial decay

Fixed only:

```{r}

modprecm_NC<-lm(k_value_tr ~prec+N_pc+C_pc,
             data=filter(mmm,termite_exposure ==0)
)
model_parameters(modprecm_NC)

```

Mixed:

```{r}

modprecm_NC<-lmer(k_value_tr ~prec+N_pc+C_pc + (1|site),
             data=filter(mm,termite_exposure ==0)
)
model_parameters(modprecm_NC, effects='fixed')

```


# Table S8. Best fit multivariate model for probably of termite discovery versus climatic sensitivities and wood chemistry

Fixed- and mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv")  %>%
  mutate(abs_lat = abs(lat),
         date_diff_years = date_diff/365) %>%
  filter(trt %in% c("T") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara")

```

Fixed only: 

```{r}

modinteraction_stoic<-glm(termite_exposure ~ prec*temp+N_pc+C_pc,
                          data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))
model_parameters(modinteraction_stoic,exponentiate = TRUE)

```

Mixed:

```{r}

modinteraction_stoic<-glmer(termite_exposure ~ prec*temp+N_pc+C_pc + (1|site),
                          data=filter(mm, trt == 'T'), offset=date_diff_years, family=binomial(link = "logit"))
model_parameters(modinteraction_stoic,exponentiate = TRUE, effects='fixed')

```


# Table S9. Best fit multivariate model for microbe (termite undiscovered) wood decay (k) versus climatic sensitivities and wood chemistry

Fixed-effects models are fitted after calculating site-level means for decay and weighting sites by the number of observations in each category. Mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv") %>%
  filter(trt %in% c("T", "C") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara") %>%
  mutate(date_diff_years = date_diff/365, 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))


#get site level decay
mm %>% 
  group_by(site, termite_exposure) %>% 
  summarise(temp = temp[1], 
            prec = prec[1], 
            forestcover100 = forestcover100[1], 
            alt = alt[1],
            date_diff_years=date_diff_years[1],
            lat = lat[1],
            N_pc = N_pc[1],
            C_pc = C_pc[1],
            n = n(), 
            k_value= mean(k_value),
            #kval = log(k_value),
            discovered = mean(termite_exposure)) %>%
  ungroup() %>% 
  mutate(termite_exposure = factor(termite_exposure), 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))->mmm

```

Fixed only: 

```{r}

m1w_NC <- lm(k_value_tr ~ temp*prec_m+N_pc+C_pc ,
             data=filter(mmm,termite_exposure ==0), weights=n)
model_parameters(m1w_NC)

```

Mixed:

```{r}

m1w_NC <- lmer(k_value_tr ~ temp*prec_m+N_pc+C_pc + (1|site),
             data=filter(mm,termite_exposure ==0))
model_parameters(m1w_NC, effects='fixed')

```


# Table S10. Best fit multivariate model for microbe + termite (termite discovered) wood decay (k) versus climatic sensitivities and wood chemistry

Fixed-effects models are fitted after calculating site-level means for decay and weighting sites by the number of observations in each category. Mixed-effects models are fitted for individual-level measurements with site-level random effects.

```{r}

mm <- read_csv("data/decomp_with_covar.csv") %>%
  filter(trt %in% c("T", "C") &
           !is.na(termite_exposure) &
           species != "Simarouba_amara") %>%
  mutate(date_diff_years = date_diff/365, 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))


#get site level decay
mm %>% 
  group_by(site, termite_exposure) %>% 
  summarise(temp = temp[1], 
            prec = prec[1], 
            forestcover100 = forestcover100[1], 
            alt = alt[1],
            date_diff_years=date_diff_years[1],
            lat = lat[1],
            N_pc = N_pc[1],
            C_pc = C_pc[1],
            n = n(), 
            k_value= mean(k_value),
            #kval = log(k_value),
            discovered = mean(termite_exposure)) %>%
  ungroup() %>% 
  mutate(termite_exposure = factor(termite_exposure), 
         abs_lat = abs(lat), 
         prec_m = prec/1000, 
         k_value_tr = log(k_value))->mmm

```

Fixed only: 

```{r}

m2w_NC <- lm(k_value_tr ~ temp*prec_m+N_pc+C_pc ,
             data=filter(mmm,termite_exposure ==1), weights=n)
model_parameters(m2w_NC)

```

Mixed:

```{r}

m2w_NC <- lmer(k_value_tr ~ temp*prec_m+N_pc+C_pc + (1|site),
             data=filter(mm,termite_exposure ==1))
model_parameters(m2w_NC, effects='fixed')

```

